\Large\bf FlexibleSUSY --- A spectrum generator generator for supersymmetric models

### 2014-11-06 doc/flexiblesusy-paper.tex
explain the manual setting of EWSB solutions in more detail

\fs assumes that each SUSY model contains Higgs bosons, which trigger
a spontaneous breaking of the electroweak symmetry.  The corresponding
EWSB consistency conditions are formulated in \fs at the one-loop
level as
%
\begin{align}
  0 = \frac{\partial V^\text{tree}}{\partial v_i} - t_i,
  \label{eq:one-loop-ewsb-eq}
\end{align}
%
where $V^\text{tree}$ is the tree-level Higgs potential, $v_i$ is the
VEV corresponding to the Higgs fields $H_i$ and $t_i$ is the one-loop
tadpole diagram of $H_i$.  Already at the tree-level ($t_i = 0$) Eqs.\
\eqref{eq:one-loop-ewsb-eq} can have multiple solutions, depending on
which paramters are chosen to be fixed by these equations.  For
example in the real MSSM Eqs.\ \eqref{eq:one-loop-ewsb-eq} can only
determine $|\mu|$, while the sign of $\mu$ is not fixed.  Currently
\fs handles only real model parameters and renormalizable theories,
which restricts the number of possible multiple solutions of Eqs.\
\eqref{eq:one-loop-ewsb-eq} to be less or equal to four times the
number of equations.  In some cases, as for example in the real MSSM,
% where $\mu$ is chosen to be fixed by the EWSB equations
these different solutions might be related by one or more free global
signs in the form
%
\begin{align}
  \mu = \sign\mu \cdot \sqrt{\ldots} .
\end{align}
%
\fs recognizes such cases and introduces these global signs as
additional free parameters.  The case of the real MSSM, where only
$|\mu|$ is fixed by the EWSB equations, is therefore handled
automatically.  If, however, the different solutions are not related
by global signs, then \fs writes all solutions to the file
\code{models/<model>/<model>_tree_level_EWSB_solution.m}.  The user
has then the option to pick a particular solution by setting it in the
\code{TreeLevelEWSBSolution} variable in the model file.  This is for
example done in the $Z_3$-violating NMSSM (\code{SMSSM}), where $\mu$
is chosen to be fixed by the EWSB equations: There the tree-level
solution for $\mu$ has the form
%
\begin{align}
  \mu = -\frac{v_s \lambda}{\sqrt{2}} + \sign X \cdot \sqrt{\ldots}
\end{align}
%
with a non-global $\sign X$.  In this model the user can parametrize
the tree-level solution in terms of $\sign X$ in the
\code{TreeLevelEWSBSolution} variable and add $\sign X$ by hand to the
list of free input parameters, see
\code{model_files/SMSSM/FlexibleSUSY.m.in} for an example.  If the
user decides to not pick a particular solution by leaving the variable
\code{TreeLevelEWSBSolution} empty, \fs tries to find a solution to
the tree-level EWSB equations numerically via an iteration.  In this
case, however, the user does not have the option to chose between the
different solutions.

>>>>

\fs assumes that each SUSY model contains Higgs bosons, which trigger
a spontaneous breaking of the electroweak symmetry.  The corresponding
EWSB consistency conditions are formulated in \fs at the one-loop
level as
%
\begin{align}
  0 &= \frac{\partial V^\text{tree}}{\partial v_i} - t_i,
  & & (i=1,\ldots,N)
  \label{eq:one-loop-ewsb-eq}
\end{align}
%
where $V^\text{tree}$ is the tree-level Higgs potential, $v_i$ is the
VEV corresponding to the Higgs fields $H_i$ and $t_i$ is the one-loop
tadpole diagram of $H_i$.  Already at the tree-level ($t_i = 0$) Eqs.\
\eqref{eq:one-loop-ewsb-eq} can have multiple solutions, depending on
which paramters are chosen to be fixed by these equations.  A
well-known example is the real MSSM, where the $\mu$-parameter is
chosen to be fixed by the EWSB equations.  In this model, the EWSB
equations, Eqs.\ \eqref{eq:one-loop-ewsb-eq} can only determine
$|\mu|$, while the sign of $\mu$ is not fixed.  This results in two
solutions for $\mu$ of the form
%
\begin{align}
  \mu = \sign\mu \cdot |\mu| .
  \label{eq:EWSB-solution-for-mu}
\end{align}
%
Currently, \fs handles only real model parameters and renormalizable
theories, which restricts the number of possible multiple solutions of
Eqs.\ \eqref{eq:one-loop-ewsb-eq} to be less or equal to $4N$.  In
some cases, as for example in the real MSSM, the different solutions
are related by one or more global signs as in
Eq.~\eqref{eq:EWSB-solution-for-mu}.  \fs recognizes such cases and
introduces these global signs as additional free parameters to allow
the user to chose between the different solutions.  The case of the
real MSSM Eq.~\eqref{eq:EWSB-solution-for-mu} is therefore handled
automatically in \fs, because $\sign\mu$ is automatically introduced
as an additional free parameter.

If, however, the different solutions are not related by global signs,
then \fs writes all solutions to the file
\code{models/<model>/<model>_tree_level_EWSB_solution.m}.  The user
has then the option to pick a particular solution by setting it in the
\code{TreeLevelEWSBSolution} variable in the model file.  This is for
example the case in the $Z_3$-violating NMSSM (\code{SMSSM}), where
$\mu$ is chosen to be fixed by the EWSB equations: There the
tree-level solution for $\mu$ has the form
%
\begin{align}
  \mu = -\frac{v_s \lambda}{\sqrt{2}} + \sign X \cdot \sqrt{\ldots}
\end{align}
%
with a non-global $\sign X$.  When \fs solves the tree-level EWSB
equations for $\mu$, $B\mu$ and $\xi_S$ it finds the two solutions and
writes them to \code{models/SMSSM/SMSSM_tree_level_EWSB_solution.m} in
the form
%
\begin{lstlisting}[language=Mathematica]
{
   {{B[\[Mu]] -> ...}},
   {
      {\[Mu] -> (-20*Sqrt[2]*vd*vS*(\[Lambda] + conj[\[Lambda]])
                 - Sqrt[...])/(80*vd)},
      {\[Mu] -> (-20*Sqrt[2]*vd*vS*(\[Lambda] + conj[\[Lambda]])
                 + Sqrt[...])/(80*vd)}
   },
   {{L[L1] -> ...}}
}
\end{lstlisting}
%
In the above solutions the dots stand for the full expression, which
is left out here for better readability.  Inspecting the two above
solutions for $\mu$, one finds that both can be parametrized by an
additional free sign in front of the \code{Sqrt[...]}.  The user can
now introduce an additional free sign by hand in the \code{MINPAR}
block
%
\begin{lstlisting}[language=Mathematica]
MINPAR = {
   {1, m0},
   {2, m12},
   {3, TanBeta},
   {4, Sign[X]},  (* <-- additional free sign *)
   {5, Azero}
};
\end{lstlisting}
%
and set the the tree-level solution, parametrized in terms of
\code{Sign[X]}, in the \code{TreeLevelEWSBSolution} variable in
the model file:
%
\begin{lstlisting}
TreeLevelEWSBSolution = {
   { B[\[Mu]], ... },
   { \[Mu], (-20*Sqrt[2]*vd*vS*(\[Lambda] + conj[\[Lambda]])
             + Sign[X] * Sqrt[...])/(80*vd) },
   { L[L1], ... }
};
\end{lstlisting}
%
One can now chose between the two solutions by setting entry $4$ in
the \code{MINPAR} block of the SLHA input file to either $+1$ or $-1$.
See \code{model_files/SMSSM/FlexibleSUSY.m.in} for a complete example
model file.  If the user decides to not pick a particular solution by
leaving the variable \code{TreeLevelEWSBSolution} empty, \fs tries to
find a solution to the tree-level EWSB equations numerically via an
iteration.  In this case, however, the user does not have the option
to chose between the different solutions.
